---
title: "Salinity and microbial inocula effects on marsh cordgrass productivity (field data)"
author: "Kai Davis, Mary-Margaret McKinney, Rachel K. Gittman, Ariane L. Peralta"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  fig_caption: yes
  pdf_document: null
header-includes:
- \usepackage{array}
- \usepackage{graphics}
- \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) #clears history

# Set working directory:
setwd("~/GitHub/2023_PSM_SMM_microbiomes/analyses")

# Standard error (se) and confidence interval (ci):
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code dependencies:
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan"); require("reshape"); require("ggplot2"): require("tidyr"); require("dplyr"); require("dbstats"); require("reshape2"); require("ggpubr"); require("glue"); require("tidyverse"); require("labdsv"); require("phyloseq")
```

# Load files
*READ IN DESIGN FILE*
```{r Import Files - design, include=FALSE}
file_path <- ("../data/PSM_SMM_design.csv")
design <- read.csv(file_path, row.names=1)
dim(design) #117  11
str(design)

#Convert chr to factors 
design$Study <- as.factor(design$Study)
design$Microbe_TRT <- as.factor(design$Microbe_TRT)
design$Water_TRT <- as.factor(design$Water_TRT)
design$Date2 <- as.factor(design$Date2)
design$TRT_row <- as.factor(design$TRT_row)
design$Source <- as.factor(design$Source) # coded as categorical for bulk vs rhiz
str(design)
```

#BACTERIA
*READ IN BACTERIAL SEQUENCING FILES*
```{r Import Files - Bacteria, include=FALSE}
#READ IN OTU FILE

otu_file <-("../data/PSM_SMM.trim.contigs.good.unique.good.filter.unique.precluster.denovo.vsearch.pick.opti_mcc.shared")
otu <- read.otu(otu_file)

#READ IN TAXONOMY FILE

tax_file <- ("../data/PSM_SMM.trim.contigs.good.unique.good.filter.unique.precluster.denovo.vsearch.pick.opti_mcc.0.03.cons.taxonomy")
tax <- import_mothur(mothur_constaxonomy_file =  tax_file)
colnames(tax)=c("Domain","Phylum","Class", "Order","Family","Genus")
tax_df <- as.data.frame(tax)

otu.tax <- read.tax(taxonomy = "../data/PSM_SMM.trim.contigs.good.unique.good.filter.unique.precluster.denovo.vsearch.pick.opti_mcc.0.03.cons.taxonomy",
                   format = "rdp", tax.levels = 6, col.tax = 3)
tax_df <- as.data.frame(otu.tax)
```

*REMOVING SINGLETONS* 
```{r remove singletons, include=FALSE}
dim(otu) #117 71805

# OTU table - remove otus w/ < 2 occurrences across all sites
otu.2 <- otu[, which(colSums(otu) > 2)]
dim(otu.2) #117 38970

# OTU table - remove otus w/ < 10 occurrences across all sites
otu.10 <- otu[, which(colSums(otu) > 10)]
dim(otu.10) #117 20446
``` 
*TRANSFORM DATA*
```{r transform data, include=FALSE}
# Correct Sample IDs and Subset File
#check number of seqs per sample
#otu_sample <- rowSums(otu.2)
otu_sample <- rowSums(otu.10)
otu_sample

# OTU table - removed low abundance samples
otu_keep <- otu.10[which(rowSums(otu.10) >= 10000), ] #removed temporarily PSM_65 b/c it represents low reads from autoclaved field soil; SMM_08 failed PCR
dim(otu_keep) #removed three samples 

# OTU table - odd sites in bacterial composition data and remove in design file and remove positive control and remove all baseline samples
odd.sites <- c("Mock_Pos","PSM_65", "SMM_08", "PSM_61", "PSM_62","PSM_63", "PSM_64", "PSM_66", "PSM_67", "PSM_68")

otu_final <- otu_keep[setdiff(rownames(otu_keep), odd.sites), ]
design_final <- design[setdiff(rownames(design), odd.sites), ]
all.equal(rownames(design_final), rownames(otu_final))

dim(design_final) #107  11
design_PSM <- subset(design_final, Study == "PSM")
dim(design_PSM) #54 11
design_SMM <- subset(design_final, Study == "SMM")
dim(design_SMM) #53 11

otu_PSM <- otu_final[c(1:54),]
dim(otu_PSM) #54 20446
all.equal(rownames(design_PSM), rownames(otu_PSM))

otu_SMM <- otu_final[c(55:107),]
dim(otu_SMM) #53 20446
all.equal(rownames(design_SMM), rownames(otu_SMM))

#now that PSM and SMM are separated, we will apply removing rare OTUs that might have been missed
otu_PSM.10 <- otu_PSM[, which(colSums(otu_PSM) > 10)]
dim(otu_PSM.10) # 54 8331

otu_SMM.10 <- otu_SMM[, which(colSums(otu_SMM) > 10)]
dim(otu_SMM.10) # 53 14812

#RELATIVE ABUNDANCE
otu_rel_PSM <- otu_PSM.10
for(i in 1:dim(otu_PSM.10)[1]){
  otu_rel_PSM[i,] <- otu_PSM.10[i,]/sum(otu_PSM.10[i,])
}

otu_rel_SMM <- otu_SMM.10
for(i in 1:dim(otu_SMM.10)[1]){
  otu_rel_SMM[i,] <- otu_SMM.10[i,]/sum(otu_SMM.10[i,])
}
```

*ADD DIVERSITY METRICS*
```{r Alpha diversity bacteria - SMM, include=FALSE}
library(vegan)

min(rowSums(otu_SMM.10)) #20108

otus.r <- rrarefy(otu_SMM.10, 20108)

# Fisher's Alpha
fisher <- fisher.alpha(otus.r)

# Species Richness
richness <- rowSums((otu_SMM.10 >= 1))

# Shannon Diversity
shannon <- vegan::diversity(otus.r, index = "shannon")
#shannon <- diversity(otus.r, index = "shannon") - error here

# Simpson's Evenness
simp.even <- apply(otus.r, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(otus.r[,-c(1:1)]))

diversity_SMM <- cbind(design_SMM,richness,shannon,simp.even,J)
#write.csv(diversity,"../data/2014to2020.diversity.bact.raw.csv")

str(diversity_SMM) 
dim(diversity_SMM) #53 15
```

#Plot Shannon diversity - SMM
```{r Plot - Shannon Diversity SMM - field, echo=TRUE}
str(diversity_SMM)
levels(diversity_SMM$Date2)
#[1] "1_June12" "2_July13" "3_Oct5" 

# Graphing Shannon diversity
p <- ggplot(diversity_SMM, aes(x=TRT_row, y=shannon, color=TRT_row, shape=Source)) +
  geom_point(aes(color=TRT_row), size=3, position = position_jitterdodge()) 
p1 <- p +facet_wrap(~Date2)

shannon <- p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",linewidth =1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Row location (decreasing elevation)", y = "Shannon Diversity Index (H')") +
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", linewidth =1)) + theme(axis.text.x = element_text(angle = )) +
    labs(color = "Row location", shape = "Source")
shannon
ggsave("../figures/SMM.shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```


#Plot Pielou's Evenness - SMM
```{r Plot - evenness - SMM, echo=TRUE}
# Graphing evenness
p <- ggplot(diversity_SMM, aes(x=TRT_row, y=J, color=TRT_row, shape=Source)) +
  geom_point(aes(color=TRT_row), size=3, position = position_jitterdodge()) 
p1 <- p +facet_wrap(~Date2)

evenness <- p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",linewidth=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Row location (decreasing elevation)", y = "Pielou's Evenness (J')") +
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", linewidth=1)) + theme(axis.text.x = element_text(angle = )) +
    labs(color = "Row location", shape = "Source")
evenness
ggsave("../figures/SMM.evenness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

## PERMANOVA Permutational multivariate analysis of variance (PERMANOVA):
```{r PERMANOVA-SMM}
otu_design_SMM <- cbind(design_SMM,otu_rel_SMM)
dim(otu_design_SMM) #53 14823
dim(design_SMM) #53 11
otu_design_SMM <- droplevels(otu_design_SMM)

adonis_SMM_row <- adonis2(formula=otu_design_SMM[,-c(1:11)]~TRT_row*Source, 
                      method = "bray", 
                      data = otu_design_SMM, 
                      perm=1000,
                      by = "terms",
                      set.seed=42)
adonis_SMM_row
```

## PCoA - average
```{r PCoA: by source}
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate
SMM_BC <- vegdist(otu_design_SMM[,-c(1:11)], method="bray")

pcoa_SMM <- cmdscale(SMM_BC, k=2, eig=TRUE, add=TRUE) # changing k to 2 because we only want 2 dimensions anyway
  # Now trying add = TRUE because of https://www.youtube.com/watch?v=G5Qckqq5Erw

explainvar1b_SMM <- round(pcoa_SMM$eig[1] / sum(pcoa_SMM$eig), 3) * 100
explainvar2b_SMM <- round(pcoa_SMM$eig[2] / sum(pcoa_SMM$eig), 3) * 100
sum.eigb <- sum(explainvar1b_SMM, explainvar2b_SMM)

explainvar1b_SMM # 34.4
explainvar2b_SMM # 11.1

pcoa.groups <- paste(otu_design_SMM$TRT_row, otu_design_SMM$Source,sep = "_")
pcoa.points <- data.frame(pcoa_SMM$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
row <- c("1","1","2","2","3","3","4","4","5","5","6","6","7","7","8","8","9","9")
source <- c("bulk","rhiz","bulk","rhiz","bulk","rhiz","bulk","rhiz","bulk","rhiz","bulk","rhiz","bulk","rhiz","bulk","rhiz","bulk","rhiz")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$row <- as.factor(row)
pcoa.cent.dataframe.trts$source <- as.factor(source)
view(pcoa.cent.dataframe.trts)
df1a <- pcoa.cent.dataframe.trts

#Plot
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=row, shape = source)) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  geom_point(aes(colour=row), size=5, stroke = 1.25, show.legend = TRUE) +
  scale_shape_manual(labels = c("bulk","rhizosphere"), values = c(16,17)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", linewidth = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (34.4%)") + ylab("PCoA 2 (11.1%)") +
        labs(colour = "Row location", shape = "Source") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    ggtitle("Bacterial 16SrRNA Community Composition")

ggsave("../figures/SMM.bact.ordination.mean.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

## PCoA - SMM indiv
```{r PCoA - SMM indiv}
#SMM_BC <- vegdist(otu_design_SMM[,-c(1:11)], method="bray")

#pcoa_SMM <- cmdscale(SMM_BC, k=2, eig=TRUE, add=TRUE) 

#explainvar1b_SMM <- round(pcoa_SMM$eig[1] / sum(pcoa_SMM$eig), 3) * 100
#explainvar2b_SMM <- round(pcoa_SMM$eig[2] / sum(pcoa_SMM$eig), 3) * 100
#sum.eigb <- sum(explainvar1b_SMM, explainvar2b_SMM)

#explainvar1b_SMM # 34.4
#explainvar2b_SMM # 11.1

pcoa_SMM.groups <- paste(otu_design_SMM$TRT_row, otu_design_SMM$Source, sep = "_")
pcoa_SMM.points <- data.frame(pcoa_SMM$points, group = pcoa_SMM.groups)

pcoa.SMM.cent.dataframe.trts <- as.data.frame(pcoa_SMM.points)

# Adding columns with the factors. H: hydrologic history
row <- c("1", "2","3","4","5","6","7","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9") 
source <- c("bulk", "bulk","bulk","bulk","bulk","bulk","bulk","bulk","rhiz", "rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","bulk", "bulk","bulk","bulk","bulk","bulk","bulk","bulk","bulk","rhiz", "rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","bulk", "bulk","bulk","bulk","bulk","bulk","bulk","bulk","bulk","rhiz", "rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz")

pcoa_SMM_indiv <- cbind(pcoa.SMM.cent.dataframe.trts,row, source)
str(pcoa_SMM_indiv)
pcoa_SMM_indiv$row <- as.factor(pcoa_SMM_indiv$row)
pcoa_SMM_indiv$source <- as.factor(pcoa_SMM_indiv$source)
dim(pcoa_SMM_indiv)
df1a <- pcoa_SMM_indiv
str(df1a)

#Plot using ggplot2 - individual no Date2

bact_SMM_indiv<- ggplot(df1a, aes(x=X1, y=X2, shape=source, colour=row)) +   
  geom_point(aes(shape = source, colour = row), size = 4) +
#removes gridlines from plot
theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
axis.line = element_line(colour = "black")) +
theme(axis.title = element_text(size=12),# face="bold"),
axis.text.x = element_text(size=14, color="black"),  axis.text.y = element_text(size=14, color="black"),
panel.border = element_rect(colour = "black", linewidth =1.25)) +
#Set plot title text size
theme(plot.title=element_text(size=12)) +
#Set legend text size
theme(legend.text=element_text(size=10, face="bold"), legend.title = element_text(size=10, face="bold"))+
#Sets size of tick marks on axis
theme(axis.ticks.length=unit(0.3,"cm")) +
#Sets labels for plot title, axis titles, and legend headings
 xlab("PCoA 1 (34.4%)") + ylab("PCoA 2 (11.1%)") +
scale_shape_manual(labels = c("bulk","rhizosphere"), values = c(16,17)) +
labs(shape = "Source") +
labs(colour="Row location") + 
guides(color = guide_legend(override.aes = list(size = 1))) +
ggtitle("Hammocks Beach Bacterial 16SrRNA Community Composition")

bact_SMM_indiv
ggsave("../figures/SMM.bact.indiv.ordination.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

## PCoA - SMM indiv - rows 1,5,9
```{r PCoA - SMM individual rows 1, eval=FALSE, include=FALSE}
pcoa_SMM.points_row159 <- pcoa_SMM_indiv[pcoa_SMM_indiv$row %in% c("1", "5", "9"), ]

source_Date2 <-c(
  "1_June12_bulk","1_June12_bulk","1_June12_bulk","1_June12_rhiz","1_June12_rhiz","1_June12_rhiz", 
  "2_July13_bulk","2_July13_bulk","2_July13_bulk","2_July13_rhiz","2_July13_rhiz","2_July13_rhiz",
  "3_Oct5_bulk", "3_Oct5_bulk", "3_Oct5_bulk", "3_Oct5_rhiz", "3_Oct5_rhiz", "3_Oct5_rhiz")

df1a <- cbind(pcoa_SMM.points_row159,source_Date2)
df1a$source_Date2 <- as.factor(df1a$source_Date2)
str(df1a)

#Plot using ggplot2 - individual no Date2

bact_SMM_indiv_159 <- ggplot(df1a, aes(x=X1, y=X2, shape=source_Date2, colour=row)) +   
  geom_point(aes(shape = source_Date2, colour = row), size = 4, stroke=1) +
  scale_shape_manual(labels = c("1_June12_bulk","1_June12_rhiz","2_July13_bulk", "2_July13_rhiz", "3_Oct5_bulk", "3_Oct5_rhiz"), values = c(15, 0, 16, 1, 17, 2)) +
#removes gridlines from plot
theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
axis.line = element_line(colour = "black")) +
theme(axis.title = element_text(size=12),# face="bold"),
axis.text.x = element_text(size=14, color="black"),  axis.text.y = element_text(size=14, color="black"),
panel.border = element_rect(colour = "black", linewidth =1.25)) +
#Set plot title text size
theme(plot.title=element_text(size=12)) +
#Set legend text size
theme(legend.text=element_text(size=10, face="bold"), legend.title = element_text(size=10, face="bold"))+
#Sets size of tick marks on axis
theme(axis.ticks.length=unit(0.3,"cm")) +
#Sets labels for plot title, axis titles, and legend headings
 xlab("PCoA 1 (34.4%)") + ylab("PCoA 2 (11.1%)") +
labs(shape = "Source") +
labs(colour="Row location") + 
scale_colour_manual(labels = c("landward", "midpoint","seaward"),
                    values = c("#D81B60","#1E88E5", "#8D82C6")) +
guides(color = guide_legend(override.aes = list(size = 4))) + 
ggtitle("Hammocks Beach Bacterial 16SrRNA Community Composition")

bact_SMM_indiv_159
ggsave("../figures/SMM.bact.indiv.159.ordination.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
## PCoA - SMM indiv Date2
```{r PCoA - SMM indiv Date2}
#SMM_BC <- vegdist(otu_design_SMM[,-c(1:11)], method="bray")
#df_SMM_BC <- as.data.frame(SMM_BC)
#pcoa_SMM <- cmdscale(SMM_BC, k=2, eig=TRUE, add=TRUE) 

#explainvar1b_SMM <- round(pcoa_SMM$eig[1] / sum(pcoa_SMM$eig), 3) * 100
#explainvar2b_SMM <- round(pcoa_SMM$eig[2] / sum(pcoa_SMM$eig), 3) * 100
#sum.eigb <- sum(explainvar1b_SMM, explainvar2b_SMM)

#explainvar1b_SMM # 34.4
#explainvar2b_SMM # 11.1

pcoa_SMM.groups <- paste(otu_design_SMM$TRT_row, otu_design_SMM$Source, otu_design_SMM$Date2, sep = "_")
pcoa_SMM.points <- data.frame(pcoa_SMM$points, group = pcoa_SMM.groups)

pcoa.SMM.cent.dataframe.trts <- as.data.frame(pcoa_SMM.points)

# Adding columns with the factors.
row <- c("1", "2","3","4","5","6","7","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9","1", "2","3","4","5","6","7","8","9") 

source <- c("bulk", "bulk","bulk","bulk","bulk","bulk","bulk","bulk","rhiz", "rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","bulk", "bulk","bulk","bulk","bulk","bulk","bulk","bulk","bulk","rhiz", "rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","bulk", "bulk","bulk","bulk","bulk","bulk","bulk","bulk","bulk","rhiz", "rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz","rhiz")

source_Date2 <-c("1_June12_bulk","1_June12_bulk","1_June12_bulk","1_June12_bulk","1_June12_bulk","1_June12_bulk","1_June12_bulk","1_June12_bulk","1_June12_rhiz", "1_June12_rhiz","1_June12_rhiz","1_June12_rhiz","1_June12_rhiz","1_June12_rhiz","1_June12_rhiz","1_June12_rhiz","1_June12_rhiz","2_July13_bulk", "2_July13_bulk","2_July13_bulk","2_July13_bulk","2_July13_bulk","2_July13_bulk","2_July13_bulk","2_July13_bulk","2_July13_bulk","2_July13_rhiz", "2_July13_rhiz","2_July13_rhiz","2_July13_rhiz","2_July13_rhiz","2_July13_rhiz","2_July13_rhiz","2_July13_rhiz","2_July13_rhiz","3_Oct5_bulk", "3_Oct5_bulk","3_Oct5_bulk","3_Oct5_bulk","3_Oct5_bulk","3_Oct5_bulk","3_Oct5_bulk","3_Oct5_bulk","3_Oct5_bulk","3_Oct5_rhiz", "3_Oct5_rhiz","3_Oct5_rhiz","3_Oct5_rhiz","3_Oct5_rhiz","3_Oct5_rhiz","3_Oct5_rhiz","3_Oct5_rhiz","3_Oct5_rhiz")
  
pcoa_SMM_indiv_Date2 <- cbind(pcoa.SMM.cent.dataframe.trts,row, source, source_Date2)
str(pcoa_SMM_indiv_Date2)
pcoa_SMM_indiv_Date2$row <- as.factor(pcoa_SMM_indiv_Date2$row)
pcoa_SMM_indiv_Date2$source <- as.factor(pcoa_SMM_indiv_Date2$source)
pcoa_SMM_indiv_Date2$source_Date2 <- as.factor(pcoa_SMM_indiv_Date2$source_Date2)
dim(pcoa_SMM_indiv_Date2)
df2a <- pcoa_SMM_indiv_Date2
str(df2a)

#Plot using ggplot2 - individual with Date2

bact_SMM_indiv_Date2<- ggplot(df2a, aes(x=X1, y=X2, shape=source_Date2, colour=row)) +   
  geom_point(aes(shape = source_Date2, colour = row), size = 4) +
  scale_shape_manual(labels = c("1_June12_bulk","1_June12_rhiz","2_July13_bulk", "2_July13_rhiz", "3_Oct5_bulk", "3_Oct5_rhiz"),
                     values = c(15, 0, 16, 1, 17, 2)) +
#removes gridlines from plot
theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
axis.line = element_line(colour = "black")) +
theme(axis.title = element_text(size=12),# face="bold"),
axis.text.x = element_text(size=14, color="black"),  axis.text.y = element_text(size=14, color="black"),
panel.border = element_rect(colour = "black", linewidth =1.25)) +
#Set plot title text size
theme(plot.title=element_text(size=12)) +
#Set legend text size
theme(legend.text=element_text(size=10, face="bold"), legend.title = element_text(size=10, face="bold"))+
#Sets size of tick marks on axis
theme(axis.ticks.length=unit(0.3,"cm")) +
#Sets labels for plot title, axis titles, and legend headings
 xlab("PCoA 1 (34.1%)") + ylab("PCoA 2 (11.0%)") +
labs(shape = "Source") +
labs(colour="Row location") + 
guides(color = guide_legend(override.aes = list(size = 2))) +
ggtitle("Hammocks Beach Bacterial 16SrRNA Community Composition")

bact_SMM_indiv_Date2
ggsave("../figures/SMM.bact.indiv.Date.ordination.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```


# Bacterial community indicator species analysis
```{r Bacteria Indicator Species i, eval=FALSE, include=FALSE}
#trying indicp spp using different package and function https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html
require("indicspecies")

# Source
levels(otu_design_SMM$Source)
(str(otu_design_SMM))
#levels(otu_design_SMM$Date2)
group <- otu_design_SMM$Source
design.type <- group
dataREL.t <- otu_rel_SMM[, colSums(otu_rel_SMM) > 0.05]
bac.ind <- indval(dataREL.t, design.type)
levels(design.type) #[1] "bulk" "rhiz"
summary(bac.ind)

inds <- which(bac.ind$pval <= 0.05)
bac.indicators <- as.data.frame(matrix(NA, nrow = length(inds), ncol = 4))
colnames(bac.indicators) <- c("OTU", "Cluster", "IndVal", "Prob")

bac.indicators$OTU <- names(inds)
bac.indicators$Cluster <- bac.ind$maxcls[inds]
bac.indicators$IndVal <- bac.ind$indcls[inds]
bac.indicators$Prob <- bac.ind$pval[inds]

ind.tax <- tax_df[which(as.character(tax_df$OTU) %in% bac.indicators$OTU), ]
ind.tax <- ind.tax[match(ind.tax$OTU, bac.indicators$OTU), ]

indicator.bac <- cbind(bac.indicators, ind.tax[, -c(1)])

indicator.bac <- indicator.bac[order(as.numeric(indicator.bac$Cluster)), ]

table(indicator.bac$Cluster)
table(indicator.bac$Phylum)
table(indicator.bac$Cluster)

# Export Bacteria Indicator Table
write.table(indicator.bac, "../data/BacterialIndicators_Source_SMM.txt",
            sep="\t", row.names = F, quote = F)
```
